### 2014-07-22 Docs/mg/mg.tex
more MG doc
lots of MG documentation work

\section{Introduction to \maestro\ Multigrid}

\section{\mgtower}

@@

\item {\tt mm}: A mask used at coarse/fine interfaces
\end{itemize}

\section{Bottom Solvers}

There are several bottom solvers available to \maestro.  These are set
through the {\tt mg\_bottom\_solver} (MAC) and {\tt
  hg\_bottom\_solver} (HG) runtime parameters.  The allowed values
  are:

@@

@@

\subsection{Special Bottom Solver}
We have a special bottom solver type that greatly speeds up large problems
and speeds up smaller problems to a lesser degree.
In fact, you should use this bottom solver whenever possible, because I
have yet to find a case that performs worse than our standard bottom solver.
What this special solver essentially does is take the data from the coarsest 
level of the original multigrid V-cycle and copy it onto a new grid structure 
with the same number of total cells in each direction, but with a fewer number of larger 
grids.  A new V-cycle begins from this point, 
so we are essentially coarsening this ``new'' problem.
Now, the coarsest level of the multigrid V-cycle in the ``new'' problem has 
fewer cells and fewer grids as compared to the original coarsest level.

@@

To understand how this bottom solver works, the first thing you need to know
is what the grid structure of the coarsest level of your multigrid V-cycle
looks like (we need to put a description of this in the section above).
Now, figure out the size of the box you would need if you
wanted it to fit all the data on the coarsest level.  Now, figure out what
the largest integer $n$ is so that you can evenly divide the length of this box
by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
since the grid structure is not suitable for this bottom solver.

Now the code will set up a ``new'' problem, using the data at the coarsest level
of the original problem as the initial data.  The grid structure for this new
problem has the same number of cells as the coarsest level of the original problem,
but the data is copied onto a grid structure where each grid has $2^n$ cells
on each side.  The new V-cycle continues down to the new coarsest level, in
which each grid has 2 cells on each side.  If you wish to impose a limit on
the maximum value that $n$ can have, you can do so by setting 
{\tt max\_mg\_bottom\_nlevs} equal to that value.

{\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$ grids, i.e., 
there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 
multigrid V-cycle contains $12\times 12\times 12$ grids that have $2^3$ cells, so the 
entire problem domain has $24^3$ cells.  We see that $n=3$, and create a new problem 
domain with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level of the 
multigrid V-cycle for the ``new'' problem will be a $3\times 3\times 3$ block of 
$2^3$ grids.

{\bf Example 2:} A 2D problem with $96\times 384$ cells divided into $48^2$ grids, i.e., 
there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 
V-cycle contains $2\times 8$ grids that have $3^2$ cells, so the entire problem 
domain has $6\times 24$ cells.  We see that $n=1$, so the program aborts since this grid
structure is not appropriate for the fancy bottom solver.

@@

\MyArrow[1em]{start1}{end1}{colorii,line width=2pt}
\MyArrow[1em]{start2}{end2}{colorii,line width=2pt}

@@

@@

  \item {\em recursively descend V-cycle}
  \begin{itemize} 
  \item \tikzmark{end1} {\tt cc\_mg\_tower\_smoother}:

@@

  \item {\em ascend V-cycle}

  \begin{itemize}
  \item\tikzmark{end2} {\tt mg\_tower\_prolongation}: Take the solution at level $n-1$ and use it to 
     correct the solution at level $n$ by representing the data on the finer grid.  This uses
     linear reconstruction for jumps by $2\times$ and piecewise-constant otherwise.

@@

@@

\section{Multigrid Solver Tolerances}

>>>>

\section{\boxlib\ Multigrid Philosophy}

Here we describe some of the general ideas behind the \boxlib\ multigrid.

We solve MG on an AMR hierarchy, which means in places we will encounter
C-F interfaces.  The \boxlib\ MG will modify the stencil for the Laplacian
at C-F interfaces to ensure the correct solution to the Poisson equation
across these interfaces.  There is no correction step needed in this
case (this differs from \cite{ricker:2008}).

The MG solver always works with jumps of $2\times$ between levels.  In 
some codes (but not \maestro) we can have jumps of $4\times$ between
levels.  \boxlib\ uses a {\tt mini\_cycle} in these cases, effectively
inserting a multigrid level between the two AMR levels and doing a mini
V-cycle.

Support for $\Delta x \ne \Delta y \ne \Delta z$

There are two MG solvers: cell-centered and nodal.

\subsection{\mgtower}

@@

\item {\tt mm}: For cell-centered, {\tt mm} takes a direction and
   tells you whether we are at a Dirichlet or Neumann boundary, or if 
   we are skewed in that direction.

   For nodal, {\tt mm} simply tells us whether a point is Dirichlet
   or Neumann.  There is no skew in nodal.

\subsection{Stencils}

There are several different stencil types that we can use for 
the discretization.  For cell-centered, these are:
\begin{itemize}
\item {\tt CC\_CROSS\_STENCIL}: this is the standard cross-stencil---5 points 
in 2-d and 7 points in 3-d.  For cell-centered MG, this is the default, and
is usually the best choice.

\item {\tt HO\_CROSS\_STENCIL}: this is a cross-stencil that uses 9 points
in 2-d and 11-points in 3-d.  For instance, it will use $(i,j)$;
$(i\pm1,j)$; $(i\pm2,j)$; $(i,j\pm1)$; $(i,j\pm2)$.  This is
higher-order accurate that {\tt CC\_CROSS\_STENCIL}.

\item {\tt HO\_DENSE\_STENCIL}: this is a dense-stencil---it uses all the
points (including corners) in $(i\pm1,j\pm1)$, resulting in a 9-point stencil
in 2-d and 27-point stencil in 3-d.
\end{itemize}

For the nodal solver, the choices are:
\begin{itemize}
\item {\tt ND\_CROSS\_STENCIL}: this is the standard cross-stencil.

\item {\tt ND\_DENSE\_STENCIL}: this is a dense stencil, using
  all the points in $(i\pm1,j\pm1)$.  The 
  derivation of this stencil is based on finite-element ideas, defining
  basis functions on the nodes.  This is developed in 2-d in 
  \cite{almgrenBellSzymczak:1996}.

\item {\tt ND\_VATER\_STENCIL}: this is an alternate dense stencil derived
  using a similar finite-element idea as above, but a different control
  volume.
\end{itemize}
For the cell-centered solve, the coefficients for the stencil are computed
once, at the beginning of the solve.  For the nodal solver, the coefficients
are hard-coded into the smoothers.

\subsection{Smoothers}

The following smoothers are available (but not necessarily for both the
cell-centered and nodal solvers):
\begin{itemize}
\item {MG\_SMOOTHER\_GS\_RB}: a red-black Gauss-Seidel smoother

\item {MG\_SMOOTHER\_JACOBI}: a Jacobi smoother (not implemented for 
the dense nodal stencil)

\item {MG\_SMOOTHER\_MINION\_CROSS}

\item {MG\_SMOOTHER\_MINION\_FULL}

\item {MG\_SMOOTHER\_EFF\_RB}
\end{itemize}

\subsection{Cycling}

The default cycling is a V-cycle, but W-cycles and full multigrid are
supported as well.

\subsection{Bottom Solvers}

The multigrid cycling coarsens the grids as part of the solve.  When 
the coarsest grid is reached, the individual boxes that comprise that
level are coarsened as much as then can, down to $2^3$ zones.  Depending
on the distribution of sizes of the grids, it may not be possible for
everything to reach this minimum size.  At this point, the bottom
solver is invoked.  Most of these will solve the linear system
on this collection of grids directly.  There is one special bottom
solver that will define a new box encompassing all of the coarsened
grids and then put the data on fewer boxes and processors and further
coarsen the problem, again until we get as close to $2^3$ as possible.
At that point, one of the other bottom solvers will be called upon
to solve the problem.

There are several bottom solvers available in \boxlib.  For \maestro.
These are set through the {\tt mg\_bottom\_solver} (MAC/cell-centered)
and {\tt hg\_bottom\_solver} (nodal) runtime parameters.
The allowed values are:
\begin{itemize}

@@

You should use the special bottom solver (4) whenever possible, even
if it means changing your gridding strategy (as discussed below) to 
make it more efficient.

@@

\subsubsection{Special Bottom Solver}

The special solver takes the data from the coarsest level of the
original multigrid V-cycle and copies it onto a new grid structure with
the same number of total cells in each direction, but with a fewer
number of larger grids.  A new V-cycle begins from this point, so we
are essentially coarsening this ``new'' problem.  Now, the coarsest
level of the multigrid V-cycle in the ``new'' problem has fewer cells
and fewer grids as compared to the original coarsest level.

@@

To understand how this bottom solver works, the first thing you need
to know is what the grid structure of the coarsest level of your
multigrid V-cycle looks like.  Next, figure out the size of the box you
would need if you wanted it to fit all the data on the coarsest level.
Finally, figure out what the largest integer $n$ is so that you can evenly
divide the length of this box by $2^n$ in every coordinate direction.
If $n < 2$, the program will abort since the grid structure is not
suitable for this bottom solver.

The code will set up a ``new'' problem, using the data at the
coarsest level of the original problem as the initial data.  The grid
structure for this new problem has the same number of cells as the
coarsest level of the original problem, but the data is copied onto a
grid structure where each grid has $2^n$ cells on each side.  The new
V-cycle continues down to the new coarsest level, in which each grid
has 2 cells on each side.  If you wish to impose a limit on the
maximum value that $n$ can have, you can do so by setting {\tt
max\_mg\_bottom\_nlevs} equal to that value.

Some grid examples help make this clear:
\begin{itemize}
\item {\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$
grids, i.e., there is a $12\times 12\times 12$ block of $32^3$ grids.
The coarsest level of the multigrid V-cycle contains $12\times
12\times 12$ grids that have $2^3$ cells, so the entire problem domain
has $24^3$ cells.  We see that $n=3$, and create a new problem domain
with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level
of the multigrid V-cycle for the ``new'' problem will be a $3\times
3\times 3$ block of $2^3$ grids.

\item {\bf Example 2:} A 2D problem with $96\times 384$ cells divided into
$48^2$ grids, i.e., there is a $2\times 8$ block of $48^2$ grids.  The
coarsest level of the multigrid V-cycle contains $2\times 8$ grids
that have $3^2$ cells, so the entire problem domain has $6\times 24$
cells.  We see that $n=0$, so the program aborts since this grid
structure is not appropriate for the fancy bottom solver.
\end{itemize}

\section{\maestro's Multigrid Use}

@@

\maestro\ multigrid solves always involve the full AMR hierarchy.

@@

The flowchart below shows the structure of a cell-centered multigrid
solve using pure V-cycles.

@@

  \item {\em recursively descend V-cycle}
\MyArrow[1em]{start1}{end1}{colorii,line width=2pt}
  \begin{itemize} 
  \item \tikzmark{end1} {\tt cc\_mg\_tower\_smoother}:

@@

  \item {\em ascend V-cycle}
\MyArrow[1em]{start2}{end2}{colorii,line width=2pt}
  \begin{itemize}
  \item\tikzmark{end2} {\tt mg\_tower\_prolongation}: Take the solution at level $n-1$ and use it to 
     correct the solution at level $n$ by representing the data on the finer grid.  This uses
     linear reconstruction for jumps by $2\times$ and piecewise-constant otherwise.

@@

\subsection{Nodal MG}

@@

\subsection{Multigrid Solver Tolerances}
